---
title: "Applied_statistics_HW1"
author: "Ganaeva Daria"
date: "01 04 2024"
output:
  pdf_document: default
  html_document: default
---

1. Install necessary packages

```{r,  message=FALSE, warning=FALSE}
#install.packages("vegan")
# install.packages("devtools")
#devtools::install_github("gavinsimpson/ggvegan")

```

2. Upload necessary packages

```{r,  message=FALSE, warning=FALSE}
library(vegan)
library(ggvegan)
```

3. Upload the data

```{r}
data(BCI)
data(BCI.env)
```

4. Exploratory data analysis
(explore data structure; check for missing values, outliers; collect summary data; filter out low abundance and non-significant data)

```{r}
#head(BCI)
head(BCI.env)

length(which(is.na(BCI)))
length(which(is.na(BCI.env)))

boxplot(BCI) 

#summary(BCI) 
BCI_f <- BCI[colMeans(BCI)>=1]

summary(BCI.env) 
BCI.env_f <- BCI.env[,-c(3,4,6)]

```

5. Ordination of objects

```{r}

ordi <- metaMDS(BCI_f, dist = "bray", autotransform = FALSE) #stress < 0.2 - we can continue

#point coordinates; merge with environmental data
ordi_pt <- data.frame(BCI.env_f, scores(ordi, display = "sites"))

#species
ordi_sp <- data.frame(scores(ordi, display = "species"))
ordi_sp$Species <- rownames(ordi_sp)

```

6. Visualization

```{r}

# Shepherd's diagram
stressplot(ordi)

#visualize using ggplot()
gg_ordi <- ggplot() +
  geom_point(data = ordi_pt, 
             aes(x = NMDS1, y = NMDS2, colour = Habitat, 
                 shape = Stream, size = EnvHet), alpha = 0.6)
gg_ordi

#add species labels 
gg_ordi_sp <- gg_ordi +
  geom_text(data = ordi_sp, 
            aes(x = NMDS1, y = NMDS2, 
                label = Species), size=2) 

gg_ordi_sp

#take another set of env factors
gg_ordi <- ggplot() +
  geom_point(data = ordi_pt, 
             aes(x = NMDS1, y = NMDS2, colour = Habitat, 
                size = UTM.EW, shape=Stream), alpha = 0.6)

gg_ordi

gg_ordi <- ggplot() +
  geom_point(data = ordi_pt, 
             aes(x = NMDS1, y = NMDS2, colour = Habitat, 
                size = UTM.NS, shape=Stream), alpha = 0.6)

gg_ordi

```

7. Interpretation of ordination

```{r}
efit <- envfit(ordi, BCI.env_f[, -c(3)])

efit$vectors

efit$factors

#statistical significance observed for UTM.EW and Habitat (P-value <0.001)
```

8. Visualization

```{r}
pal_col <- c("red", "green", "blue", "black", "orange")
pal_sh <- c(0,1)

ordiplot(ordi, type = "n")
points(ordi, col = BCI.env_f$Habitat, pch = pal_sh[BCI.env_f$Stream])
plot(efit)

#correlation observed for UTM.EW, EnvHet and Habitat OldLow
```
9. ordisurf() and visualization

```{r}
ordisurf_UE <- ordisurf(ordi, BCI.env_f$UTM.EW, method = "REML")
ordisurf_EH <- ordisurf(ordi, BCI.env_f$EnvHet, method = "REML") 

summary(ordisurf_UE)
summary(ordisurf_EH)

#correlation for UTM.EW, with community is statistically significant; for EnvHet - not
```

10. Conclusions

We discovered the relationship between tropical tree species beta-diversity and different environmental factors. 
Correlation with statistical significance was observed for such factors as Habitat and UTM (East-Western coordinates).
